---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bis620.2022

<!-- badges: start -->
[![R-CMD-check](https://github.com/Scott-Zuo/bis620.2022/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Scott-Zuo/bis620.2022/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Scott-Zuo/bis620.2022/branch/main/graph/badge.svg)](https://app.codecov.io/gh/Scott-Zuo/bis620.2022?branch=main)
<!-- badges: end -->

The goal of bis620.2022 is to generate accelerometry plot for dataset with time and 3-dimensional coordinates.
It also includes a functionality of generating spectal signature plots.
The dataset included in the package is from UKBiobank for examples

## Installation

You can install the development version of bis620.2022 from [GitHub](https://github.com/) with: <br>
`install.packages("devtools")`
`devtools::install_github("Scott-Zuo/bis620.2022")`

## Example


```{r example}
library(bis620.2022)
## basic example code
data(ukb_accel)
ukb_accel[1:100, ] |>
      spectral_signature(take_log = TRUE) |>
      accel_plot()
```
